---
title: "Project 3 - EDA"
author: "Max Lippman, Marti Walstad, Jenna Sayle, Alaina Edwards"
format: 
  html:
    toc: true
    toc-location: left
    toc-expand: true
    code-tools: true 
    code-fold: true
    code-overflow: scroll
    self-contained: true
editor: 
  markdown: 
    wrap: 72
---

## Session Information

For the purpose of data reproducibility, we included the session info and loaded all required packages in this one location. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(require(pacman)==FALSE) install.packages("pacman")
pacman::p_load(#DataExplorer, # For explorer and cleaning data by visualizing missing values.
               gptstudio,#For coding assistance
               httr, # For downloading the data from the UCI
               tidyverse, # For data manipulation
               sjPlot, 
               corrplot, # for heatmaps
               DataExplorer, #for exploring data
               skimr, # for nice summary of data
               readxl,
               dplyr
               ) # For creating dummy variables of categorical variables automatically.)
```

```{r}
sessionInfo()
```

## Time Series Analytis 

### ISA2 Applied Business Statistics Thematic Sequence over Terms
```{r}
data = read_excel("Thematic Sequence.xlsx")
p = ggplot(data, aes(x = `Term Code`, y = `Enrolled Student Count`)) +
  geom_bar(stat = "identity", color = "steelblue", fill = "steelblue", width = 0.8) +
  xlab("Term Code") +
  ylab("Enrolled Student Count") +
  ggtitle("ISA2 Thematic Sequence Enrollment over Terms") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better visibility

p
```
### Business Analytics Minor over Terms

```{r}
data = read_excel("BA Minors.xlsx")

p = ggplot(data, aes(x = `Term Code`, y = `Enrolled Student Count`)) +
  geom_bar(stat = "identity", color = "steelblue", fill = "steelblue", width = 0.8) +
  xlab("Term Code") +
  ylab("Enrolled Student Count") +
  ggtitle("Business Analytics Minor Enrollment over Terms") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

p
```
### Business Analytics Major over Term
```{r}

```
### Count of Business Analytics Majors 

```{r}
ba_major_counts_by_term <- aggregate(Major.1 ~ Term.Code, data = merged_data, function(x) sum(x == "BA Major"))


ba_major_counts_by_term2 <- aggregate(Major.2 ~ Term.Code, data = merged_data, function(x) sum(x == "BA Major"))

ba_major_counts_by_term3 <- aggregate(Major.3 ~ Term.Code, data = merged_data, function(x) sum(x == "BA Major"))

# Assuming ba_major_counts_by_term, ba_major_counts_by_term2, ba_major_counts_by_term3 are your data frames
merged_counts <- merge(ba_major_counts_by_term, ba_major_counts_by_term2, by = "Term.Code", all = TRUE)
merged_counts <- merge(merged_counts, ba_major_counts_by_term3, by = "Term.Code", all = TRUE)

# Replace missing values with zeros
merged_counts[is.na(merged_counts)] <- 0

# Set column names for clarity
names(merged_counts) <- c("Term.Code", "BA_Major_Count_1", "BA_Major_Count_2", "BA_Major_Count_3")

# Display the result
print(merged_counts)



p <- ggplot(merged_counts, aes(x = Term.Code)) +
  geom_bar(aes(y = BA_Major_Count_1), stat = "identity", position = "dodge", fill = "steelblue", color = "steelblue", width = 5) +
  geom_bar(aes(y = BA_Major_Count_2), stat = "identity", position = "dodge", fill = "steelblue", color = "steelblue", width = 5) +
  geom_bar(aes(y = BA_Major_Count_3), stat = "identity", position = "dodge", fill = "steelblue", color = "steelblue", width = 5) +
  xlab("Term Code") +
  ylab("BA Major Count") +
  ggtitle("BA Major Counts by Term") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 8),  # Adjust the size of axis text
        panel.grid.major.x = element_blank(),  # Hide major gridlines
        panel.grid.minor.x = element_blank()) 

print(p)
```
## Random Forest Model

