---
title: "Project 3 Checkpoint 1"
author: "Max Lippman"
date: "2023-11-05"
output:
  html_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Installation

```{r}
if(require(pacman)==FALSE) install.packages("pacman")
pacman::p_load(#DataExplorer, # For explorer and cleaning data by visualizing missing values.
               gptstudio,#For coding assistance
               httr, # For downloading the data from the UCI
               tidyverse, # For data manipulation
               sjPlot, 
               corrplot, # for heatmaps
               DataExplorer, #for exploring data
               skimr # for nice summary of data
               ) # For creating dummy variables of categorical variables automatically.)
```

## Introduction

This document is meant to summarize where we currently are (group 8) in the beginning off our analysis project for ISA 616.

## Client
Skip Benamati and the Business Analytics/ISA Dept.

## Problem Statement
It is challenging to predict demand for the program and course size for the Business Analytics program. Skip and the BA department would like to know the forecasted number of students so that courses are not offered over abundantly or insufficiently.

## Research Questions we Want to Answer
* How many students are forecasted to be in the Spring 2024 BA cohort? 
* Historically, what characteristics define students who stay with the BA major?
* Historically, what characteristics define students who don’t stay with the BA major?
* What majors are people transferring to BA from?
* What courses have a significant impact on whether students stay with the BA major?

## Our Final Deliverables
* An RMarkdown document containing background information on our project scope, BVP, and questions to answer.
* A document containing our EDA analysis that guides viewers through the steps we took to clean our data.
* A dashboard containing visuals, graphs, and tables that visualize our findings per our research questions.
* A document containing our conclusions, final insights, and further recommendations to Skip and the BA department.

## Timeline
* Friday 11/3 - observe the data and define expectations for the group and project scope
Assignees : ALL

* Sunday 11/5 - have BVP submitted, Github setup (First checkpoint)
BVP Assignees : Marti
Project Documentation Assignees: Alaina, Max
GitHub repo setup: Jenna

* Thursday 11/9 - have data cleaned and merged, Rmarkdown file used that includes the documentation our work
Data Cleaning : Jenna
Data Merging: Marti 
Adding context and Relevant information to file : Max, Alaina

* Friday 11/10- Meet with client to ensure our product / model aligns with their needs
Assignees: ALL
Tuesday 11/14 - find students to potentially interview who have left or switched into the BA major (?)
Assignees: Max

* Thursday 11/16 - Dashboard template created
Assignees: Marti

* Wednesday 11/29 - final project turned in
Assignees: ALL

* 11/30, 12/5, 12/7, 12/8 - presentations to the class
Assignees: ALL

* Friday 12/8 - final project delivered to client
Assignees: ALL

## Our Business Value Proposition
